
---
title: Implementation of LISA - Variation  of Income across different LGAs in Australia 
date: May 9, 2018
output: beamer_presentation
---

#From where did we get the Data

##Income Data Source

- Australian Bureau Of Statistics (http://stat.data.abs.gov.au) 
- Economy -> Finance -> Household Income -> Census 2016, Total Household Income (Weekly) by    Household Composition (LGA)
- Customise layout():   
                       Page -  Census Year, Household Composition  
                       Row - State, Region, Geography Level  
                       Column - Total Household Income Weekly 
                       
#From where did we get the Data  

##Shapefile Data Source

- Australian Bureau Of Statistics (http://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/1270.0.55.003July%202016?OpenDocument)

- Name of the file (Zip) to download: Local Government Areas ASGS Ed 2016 Digital Boundaries in ESRI Shapefile Format 

#How does the Data look
```{r setup, include=FALSE,}
library(dplyr)
library(magrittr)
library(knitr)
library(readr)
library(ggplot2)
library(spdep)
knitr::opts_chunk$set(cache=TRUE)
```


```{r Raw Data,echo=FALSE,warnings=FALSE}
IncomeData <- read_csv('Data/Income_Data.csv',col_types = cols()) 
IncomeData[,c("Weekly_Household_Income","Annual_Household_Income","State","LGA_2016","Region","Value")]
```
#Analysis

- Number of people for each income category provided for each regions.  
- For a broad overview of the income distribution across different LGAs in Australia, weekly household income categorised into the following sections:
  
  - Low: Household Weekly Income less than 500A$
  - Medium: Household Weekly Income from 500A$ to 1500A$
  - High: Household Weekly Income Above A$1500 (ranges from A$1500 to A$8000 and beyond)

The key Average Weekly Earnings figures as provided in  
Average Weekly Earnings, Australia, Nov 2017  
(http://www.abs.gov.au/ausstats/abs@.nsf/mf/6302.0) is 1567 A$. Hence, the medium income range is defined in that range.

#Analysis

##Table2
```{r print-lga-income,echo=FALSE, dependson='show-income-data'}
Cat_Income <- IncomeData %>%
  mutate(State = case_when(
    State == 'New South Wales' ~ 'NSW',
    State == 'Victoria' ~ 'VIC',
    State == 'Queensland' ~ 'QLD',
    State == 'South Australia' ~ 'SA',
    State == 'Western Australia' ~ 'WA',
    State == 'Tasmania' ~ 'TAS',
    State == 'Northern Territory' ~ 'NT',
    State == 'Australian Capital Territory' ~ 'ACT',
    TRUE ~ as.character(State)))%>%
  mutate(
    Weekly_Household_Income = case_when(
      Weekly_Household_Income %in% c("$1-$149", "$150-$299", "$300-$399", "$400-$499") ~ 'Low',
      Weekly_Household_Income %in% c(
        "$500-$649",
        "$650-$799",
        "$800-$999",
        "$1,000-$1,249",
        "$1,250-$1,499" ) ~ 'Medium',
        Weekly_Household_Income %in% c("$1,
        500-$1,749","$1,500-$1,749",
        "$1,750-$1,999","$2,000-$2,499","$2,500-$2,999","$3,000-$3,499","$3,500-$3,999","$4,000-$4,499","$4,500-$4,999"
        ,"$5,000-$5,999","$6,000-$7,999","$8,000 or more") ~ 'High',TRUE~as.character(Weekly_Household_Income)
    ))
Cat_Income %>%
  group_by(LGA_2016,Weekly_Household_Income) %>%
  summarise(Number_People = sum(`Value`)) -> LGA_Income_catg
LGA_Income_catg

```
#Analysis

##Obtain one representative income for each LGA

Weekly household median incomes assumed for each class:  
  - Low: 250A$  
  - Medium: 1000A$  
  - High: 5000A$  



```{r print-LGA Input,echo=FALSE}

## LGA
#Reading OGR vector maps into Spatial objects
#The Geospatial Data Abstraction Library (GDAL) is a computer software library for reading and writing raster and vector geospatial data formats
#The rgdal package has been around for more than a decade and provides bindings to the incredible Geospatial Data Abstraction Library (GDAL) for reading, writing and converting between spatial formats.
LGAShp <- rgdal::readOGR('LGA_shapefiles',layer="LGA_2016_AUST")
# for topologically-aware polygon simplification
LGA_subset <- rmapshaper::ms_simplify(LGAShp, keep = 0.05)
save(LGA_subset,file = "Data/GA_subset.Rda")

#Some LGA codes not in LGAData but present in Income Data
#For which income level are available but no shape file available
#Deleting those LGA codes for the purpose of carrying out further analysis


LGA_data <- LGA_subset@data

LGA_data$id <- row.names(LGA_data)
LGA_data$LGA_CODE16 <- as.integer(as.character(LGA_data$LGA_CODE16))
LGA_data$id <- row.names(LGA_data)
LGA_data$LGA_CODE16 <- as.integer(as.character(LGA_data$LGA_CODE16))
LGA_map <- ggplot2::fortify(LGA_subset)
LGA_map$group <- paste("g",LGA_map$group,sep=".")
LGA_map$piece <- paste("p",LGA_map$piece,sep=".")
```


```{r print-Oneincome,echo=FALSE}
Income_Median <- LGA_Income_catg %>% filter(Number_People>0)%>%mutate(Median_Income = case_when(Weekly_Household_Income=='Low'~ 250,
  Weekly_Household_Income=='Medium'~1000,
  Weekly_Household_Income=='High'~5000)
)
  
  Income_Median %>%
  group_by(LGA_2016) %>%
  summarise(Median_Income_LGA = sum(`Number_People`*`Median_Income`)/sum(`Number_People`)) ->OneIncome_LGA
```
#HeatMap - Maps with Structural Breaks
```{r print-Heatmap,echo=FALSE}
LGA_data %>%
  select(id, LGA_CODE16, LGA_NAME16) %>%
  rename(name = LGA_NAME16) %>%
  right_join(OneIncome_LGA,by=c("LGA_CODE16" = "LGA_2016")) %>%
  right_join(LGA_map)-> LGA_map_Income

ggplot(LGA_map_Income) + geom_polygon(aes(long, lat, group = group,fill=Median_Income_LGA)) + coord_fixed()

```  

```{r Allcode,echo=FALSE}

#Creating k nearest Neighbours
# knn2nb converts a knn object returned by knearneigh into a neighbours list of class nb with a list of integer vectors containing neighbour region number ids.
coords <-coordinates(LGA_subset)
length(coords)
IDs <- row.names(as(LGA_subset, "data.frame"))
LGA_nn_1 <- knn2nb(knearneigh(coords, k = 1), row.names = IDs)
LGA_nn_2 <- knn2nb(knearneigh(coords, k = 2), row.names = IDs)
LGA_nn_3 <- knn2nb(knearneigh(coords, k = 3), row.names = IDs)
LGA_nn_4 <- knn2nb(knearneigh(coords, k = 4), row.names = IDs)
```

#Plotting nearest neighbors neighbours K=1

```{r,echo=FALSE}
plot(LGA_nn_1,coords)
dsts <- unlist(nbdists(LGA_nn_1, coords,longlat=TRUE))
max_1nn <- max(dsts)
max_1nn
```

#Plotting nearest neighbors neighbours K=2

```{r,echo=FALSE}
plot(LGA_nn_2,coords)
dsts <- unlist(nbdists(LGA_nn_2, coords,longlat=TRUE))
max_1nn <- max(dsts)
max_1nn
```

#Plotting nearest neighbors neighbours K=3

```{r,echo=FALSE}
plot(LGA_nn_3,coords)
dsts <- unlist(nbdists(LGA_nn_3, coords,longlat=TRUE))
max_1nn <- max(dsts)
max_1nn
```

#Plotting nearest neighbors neighbours K=4

```{r,echo=FALSE}
plot(LGA_nn_3,coords)
dsts <- unlist(nbdists(LGA_nn_4, coords,longlat=TRUE))
max_1nn <- max(dsts)
max_1nn
```

```{r,echo=FALSE}
# Checking if the neighbours object is symmetric

nb_list <- list(k1 = LGA_nn_1, k2 = LGA_nn_2, k3=LGA_nn_1, k4 = LGA_nn_1)
sapply(nb_list, function(x) is.symmetric.nb(x, verbose = FALSE,force = TRUE))
#If need be, k-nearest neighbour objects can be made symmetrical using the make.sym.nb function
LGA_nn_1 <- make.sym.nb(LGA_nn_1)

#To calculate a list of vectors of distances
#corresponding to the neighbour object for first nearest neighbours. The
#greatest value will be the minimum distance needed to make sure that all the
#areas are linked to at least one neighbour
dsts <- unlist(nbdists(LGA_nn_1, coords,longlat=TRUE))
summary(dsts)
max_1nn <- max(dsts)
max_1nn

#nb2listw function takes a neighbours list object and converts it into a weights object
#For style="W", the weights vary between unity divided by the largest and
#smallest numbers of neighbours, and the sums of weights for each areal entity
#are unity  
#B is the basic binary coding

Weights_W <- nb2listw(LGA_nn_1,style="W")
Weights_B <- nb2listw(LGA_nn_1,style="B")
Weights_C <- nb2listw(LGA_nn_1,style= "C")
Weights_U <- nb2listw(LGA_nn_1,style="U")
Weights_S <- nb2listw(LGA_nn_1,style="S")

#Weights and Neighbors same as LGA_map1.R
```

#Moran Scatterplot

```{r,echo=FALSE}
Moran_test <-moran.test(OneIncome_LGA$Median_Income_LGA, Weights_W)
moran.plot(OneIncome_LGA$Median_Income_LGA, Weights_W)
```        

#Local Moran
```{r,echo=FALSE}
#Local measures of spatial autocorrelation
local_Moran <- localmoran(OneIncome_LGA$Median_Income_LGA, Weights_W)
summary(local_Moran)
```

#Local Significance Plot using Moran's Statistic

```{r,echo=FALSE}
#creating scaled resonse variable
OneIncome_LGA$Median_Income_LGA_s <- scale(OneIncome_LGA$Median_Income_LGA)  %>% as.vector()
# create a spatially lagged variable and save it to a new column
OneIncome_LGA$Median_Income_LGA_s_lag <- scale(lag.listw(Weights_W,OneIncome_LGA$Median_Income_LGA))%>%as.vector()
# summary of variables, to inform the analysis
#summary(OneIncome_LGA$Median_Income_LGA_s)
#summary(OneIncome_LGA$Median_Income_LGA_s_lag)


OneIncome_LGA$local_Moran_stat <- NA

# high-high quadrant
OneIncome_LGA[(OneIncome_LGA$Median_Income_LGA_s >= 0 & 
                 OneIncome_LGA$Median_Income_LGA_s_lag >= 0) & 
               (local_Moran[, 5] <= 0.05), "local_Moran_stat"] <- "high-high"
# low-low quadrant

OneIncome_LGA[(OneIncome_LGA$Median_Income_LGA_s <= 0 & 
                 OneIncome_LGA$Median_Income_LGA_s_lag <= 0) & 
               (local_Moran[, 5] <= 0.05), "local_Moran_stat"] <- "low-low"

# high-low quadrant

OneIncome_LGA[(OneIncome_LGA$Median_Income_LGA_s >= 0 & 
                 OneIncome_LGA$Median_Income_LGA_s_lag  <= 0) & 
               (local_Moran[, 5] <= 0.05), "local_Moran_stat"] <- "high-low"

# low-high quadrant

OneIncome_LGA[(OneIncome_LGA$Median_Income_LGA_s <= 0 & 
                 OneIncome_LGA$Median_Income_LGA_s_lag >= 0) & 
               (local_Moran[, 5] <= 0.05), "local_Moran_stat"] <- "low-high"


# non-significant observations
OneIncome_LGA[(local_Moran[, 5] > 0.05), "local_Moran_stat"] <- "not signif."  


OneIncome_LGA %>% 
  group_by(local_Moran_stat) %>% 
  tally()
```

#Local Significance Plot using Moran's Statistic

```{r,echo=FALSE}
# plotting the map for local_Morans
df_OneIncome <- fortify(OneIncome_LGA, region="LGA_2016")
df_OneIncome %>% right_join(OneIncome_LGA) %>%right_join(LGA_data,by=c("LGA_2016"="LGA_CODE16")) %>% right_join(LGA_map)-> LGA_OneIncome

LGA_OneIncome %>% 
  ggplot(aes(long, lat, group = group, fill = local_Moran_stat)) + 
  geom_polygon(color = "white", size = .05)  + coord_equal() + 
  theme_void() + scale_fill_brewer( palette = "Set1") +coord_fixed()
```

#Local Significance Plot using Gi star Statistic

```{r,echo=FALSE,warning=FALSE}
#G and Gstar local spatial statistics
local_G <-localG(OneIncome_LGA$Median_Income_LGA, Weights_W)
summary(local_G)

local_G_new<- NA

local_G_new[local_G >= 1.96] <- "high-high"
local_G_new[(local_G <= -1.96)] <- "low-low"
local_G_new[(-1.96<local_G)& (local_G<1.96)] <- "not_signif."

LGA_OneIncome_r=cbind(OneIncome_LGA,local_G_new)

LGA_OneIncome_r %>% 
  group_by(local_G_new) %>% 
  tally()
```

#Local Significance Plot using Gi star Statistic

```{r,echo=FALSE,warning=FALSE}
# plotting the map
df_Gstar_OneIn <- fortify(LGA_OneIncome_r, region="LGA_2016")

df_Gstar_OneIn %>% right_join(OneIncome_LGA) %>%right_join(LGA_data,by=c("LGA_2016"="LGA_CODE16")) %>% right_join(LGA_map)-> LGA_OneIncome_Gstar

LGA_OneIncome_Gstar %>% 
  ggplot(aes(long, lat, group = group, fill = local_G_new)) + 
  geom_polygon(color = "white", size = .05)  + coord_equal() + 
  theme_void() + scale_fill_brewer( palette = "Set1")+ coord_fixed()
```

#Regions spatially significant in local Moran and not in local Gi star 

```{r,echo=FALSE}
#LGA_map_4 %>% filter(local_Moran_stat == c(""high-high","low-low"))
##LGA where local Moran stat is significant and high high
LGA_OneIncome_Gstar %>% 
  group_by(local_Moran_stat,LGA_NAME16,STE_NAME16) %>% filter(local_Moran_stat=="high-high") %>% tally() ->Moran_High_High

LGA_OneIncome_Gstar %>% 
  group_by(local_Moran_stat,LGA_NAME16,STE_NAME16) %>% filter(local_Moran_stat=="low-low")%>%tally()->Moran_low_low
##LGA where local  G stat is significant and high high
LGA_OneIncome_Gstar %>% 
  group_by(local_G_new,LGA_NAME16,STE_NAME16) %>% filter(local_G_new=="high-high")%>%tally()->G_High_High


LGA_OneIncome_Gstar %>% 
  group_by(local_G_new,LGA_NAME16,STE_NAME16) %>% filter(local_G_new=="low-low") %>%tally()->G_low_low

#spatially significant in local Moran but not in local Gi
Moran_not_G = setdiff(Moran_High_High$LGA_NAME16,G_High_High$LGA_NAME16)
kable(Moran_not_G,longtable=TRUE)
```

#Regions spatially significant in local Moran and not in local Gi star 

```{r,echo=FALSE}
G_not_Moran = setdiff(G_High_High$LGA_NAME16,Moran_High_High$LGA_NAME16)
kable(G_not_Moran,longtable=TRUE)
```
